\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-12-16 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R1.7. Comments by reviewer}:\\
{\em It is suggested to count the total computing time, not on each step.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.8. Comments by reviewer}:\\
{\em If the proposed method depends on the Kronecker product, how can it be scaled to a higher dimension?}

{\nib{ \blue The proposed method \ul{does not entail computation of a Kronecker product}. In the manuscript, the Kronecker product with the $2\times 2$ identity matrix appeared in the definition of the change-of-variable (17). \ul{Its only purpose is to define} that change-of-variable, and \ul{thereby derive} the PDE in the new variables. The proximal recursion (for the PDE thus derived) detailed in Algorithm 1 \ul{does not have any Kronecker product computation}.}}

@@

\noindent
{\bf R1.9. Comments by reviewer}:\\
{\em beta is undefined in equation (15a).}

{\nib{ \blue Thanks for the comment. The $\beta$ in (15a) is the same $\beta$ in (16), as in part of the same sentence. To avoid any confusion, in the revised manuscript, we have moved the phrase ``for some $\beta>0$" in (15a).}}

### 2021-12-25 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R1.6. Comments by reviewer}:\\
{\em The simulation time is too short to convince the readers. Only 1 second is considered here. Note that the errors can be quickly accumulated in the iteration.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.7. Comments by reviewer}:\\
{\em It is suggested to count the total computing time, not on each step.}

{\nib{ \blue While the depiction of the computational times in Figs. 8 and 9 in the revised manuscript are somewhat unconventional, there is a reason why we choose to plot that way. At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update involves execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. To claim that the proposed method is }}

@@

\noindent
{\bf R1.8. Comments by reviewer}:\\
{\em If the proposed method depends on the Kronecker product, how can it be scaled to a higher dimension?}

{\nib{ \blue The proposed method \ul{does not entail computation of a Kronecker product}. In the manuscript, the Kronecker product with the $2\times 2$ identity matrix appeared in the definition of the change-of-variable (17). \ul{Its only purpose is to define} that change-of-variable, and \ul{thereby derive} the PDE in the new variables. The proximal recursion (for the PDE thus derived) detailed in Algorithm 1 \ul{does not have any Kronecker product computation}.}}

\noindent
{\bf R1.9. Comments by reviewer}:\\
{\em beta is undefined in equation (15a).}

{\nib{ \blue Thanks for the comment. The $\beta$ in (15a) is the same $\beta$ in (16), as in part of the same sentence. \ul{To avoid any confusion, in the revised manuscript, we have moved the phrase ``for some $\beta>0$" in (15a)}.}}

@@

{\nib \blue{TBD.}}

\noindent
{\bf R4.3. Comments by reviewer}:\\
{\em Regarding the presentation, the reviewer suggests making the paper more self-contained. Some discussion in this paper highly relies on the authors' previous work, which can be improved by either adding a summarization or trimming out the materials.

Following the same vein, there are some unnecessary information and technical points provided in this paper that may be distracting.
For example, on page 3, why would the authors mention the Kuramoto oscillator model?
On page 4, is the detailed discussion regarding Wasserstein-2 metric necessary? Similarly, on page 5, the discussion about the natural energy functional and how the gradient drift case does not apply also seems too detailed.}

{\nib \blue{TBD FOR FIRST PARA.

We appreciate the reviewer's point that the sentence containing equation (6) and the mention of Kuramoto model can be removed without affecting the rest of the development. However, we prefer to keep it since it adds didactic value, and many engineers may feel comfortable by the explicit mention that (1) is simply a rewriting of (6). Furthermore, an astute reader, encountering that sentence will realize that the scope of the technical contribution of this paper is broad because the algorithmic framework developed here will be applicable to several other non-power system applications (mentioned in the two cited references in that sentence) where the second order Kuramoto model is used. Since this is just a matter of one sentence, we will prefer to keep it. We hope you approve.

\ul{On the other hand, we believe the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}. Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory. The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature. We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}

@@

\noindent
{\bf R4.4. Comments by reviewer}:\\
{\em Along with [25], there are recent developments that over-approximates the reachable set for nonlinear systems under uncertain initial conditions, parameters, and external disturbances:
``Reach-Set Estimation for DAE Systems under Uncertainty and Disturbances Using Trajectory Sensitivity and Logarithmic Norm".}

{\nib \blue{Thanks for providing that interesting recent reference. \ul{We have cited the same in the revised manuscript}.}}

\noindent
{\bf R4.5. Comments by reviewer}:\\
{\em On page 7, the authors stated about ``To account the pertinent geometry of $\mathbb{T}^n \times \mathbb{R}^n$". What exactly does this mean? Do the authors indicate that in (28) the two groups of state values are independent? Or do the authors mean the specific choices of the product von Mises PDF and the uniform distribution? If so, what are the connections?}

{\nib \blue{We simply mean that the joint PDF should be supported on $\mathbb{T}^n \times \mathbb{R}^n$, or a submanifold thereof. For instance, one cannot assume the initial joint PDF to be a $2n$ dimensional multivariate normal since that is supported over $\mathbb{R}^{2n}$, not over $\mathbb{T}^n \times \mathbb{R}^n$. It is, however, perfectly fine if some or all of the $2n$ components of the initial state vector are statistically correlated. As long as the support is $\mathbb{T}^n \times \mathbb{R}^n$, there is no issue if the angular variables are correlated with the Euclidean (i.e., angular velocity) variables.

}}

@@

\noindent
{\bf R4.6. Comments by reviewer}:\\
{\em On page 5, is there a mistake in ``random samples from (28), and evaluated at (28)"?}

{\nib \blue{That was not a typo. Indeed, equation (28) was \ul{used in two ways}: it was used to generate initial state samples, and then the initial joint PDF values at those samples were evaluated via the functional form of (28), i.e., as in function evaluation. This is needed in our case because unlike standard Monte Carlo, we \ul{co-evolve} the state samples and joint PDF values at those state samples. \ul{To clarify this further, in the revised manuscript, we have slightly expanded that sentence}.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R1.6. Comments by reviewer}:\\
{\em The simulation time is too short to convince the readers. Only 1 second is considered here. Note that the errors can be quickly accumulated in the iteration.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.7. Comments by reviewer}:\\
{\em It is suggested to count the total computing time, not on each step.}

{\nib{ \blue While the depiction of the computational times in Figs. 8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update involves execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time. On the other hand, theoretical consistency (see the unnumbered equation after (26), before Sec. IV) requires that $h$ itself should be small. Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear if the computational time needed to do so can be much faster than $h$. If not, then the method would be statistically consistent but practically not useful. The Figs. 8 and 9 in the revised manuscript show that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.}}

@@

Following the same vein, there are some unnecessary information and technical points provided in this paper that may be distracting.
For example, on page 3, why would the authors mention the Kuramoto oscillator model?
On page 4, is the detailed discussion regarding Wasserstein-2 metric necessary? Similarly, on page 5, the discussion about the natural energy functional and how the gradient drift case does not apply also seems too detailed.}

{\nib \blue{TBD FOR FIRST PARA.

We appreciate the reviewer's point that the sentence containing equation (6) and the mention of Kuramoto model can be removed without affecting the rest of the development. However, we prefer to keep it since it adds didactic value, and many engineers may feel comfortable by the explicit mention that (1) is simply a rewriting of (6). Furthermore, an astute reader, encountering that sentence will realize that the scope of the technical contribution of this paper is broad because the algorithmic framework developed here will be applicable to several other non-power system applications (mentioned in the two cited references in that sentence) where the second order Kuramoto model is used. Since this is just a matter of one sentence, we will prefer to keep it. We hope you approve.

\ul{On the other hand, we believe the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}. Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory. \ul{The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}. We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

{\nib{ \blue TBD.}}

\noindent
{\bf R1.7. Comments by reviewer}:\\
{\em It is suggested to count the total computing time, not on each step.}

{\nib{ \blue While the depiction of the computational times in Figs. 8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update involves execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. \ul{To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time}. On the other hand, theoretical consistency (see the unnumbered equation after (26), before Sec. IV) requires that $h$ itself should be small. Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$. If not, then the method would be statistically consistent but practically not useful. The Figs. 8 and 9 in the revised manuscript show that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.}}

@@

\noindent
{\bf R2.2. Comments by reviewer}:\\
{\em The paper's language and presentation make it seem like it is more oriented towards the applied mathematics and control communities. It is highly suggested that the authors revise the exposition of ideas and concepts, putting more emphasis on the power system so that it can be well-understood by the power system experts.}

{\nib \blue{TBD.}}

\noindent
{\bf R2.3. Comments by reviewer}:\\
{\em How will the proposed approach be used for the cases in which dependencies among uncertainty sources exist? An explanation and illustration would be helpful.}

{\nib \blue{In our paper, we do not make any assumption on statistical independence of the uncertainties. \ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R2.2. Comments by reviewer}:\\
{\em The paper's language and presentation make it seem like it is more oriented towards the applied mathematics and control communities. It is highly suggested that the authors revise the exposition of ideas and concepts, putting more emphasis on the power system so that it can be well-understood by the power system experts.}

{\nib \blue{TBD.}}

\noindent
{\bf R2.3. Comments by reviewer}:\\
{\em How will the proposed approach be used for the cases in which dependencies among uncertainty sources exist? An explanation and illustration would be helpful.}

{\nib \blue{In our paper, we do not make any assumption on statistical independence of the uncertainties. \ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}. Please note that the initial joint PDFs (28) and (30) are taken to be in that form only for the sake of concrete example simulations. Their only purpose, in the simulations reported in Sec IV, is to generate joint samples, and to evaluate the initial joint PDF values at the joint samples thus generated. If this initial scattered point cloud data is supplied to us only as a numerical table, without specifying any functional form whatsoever, nothing changes in the proposed algorithm. Similarly, the parameters in Table I could be correlated random vectors, and all we need are the \ul{joint} parametric weighted point cloud data. In fact, this is how we expect the proposed computational framework to be applied in practice.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R2.2. Comments by reviewer}:\\
{\em The paper's language and presentation make it seem like it is more oriented towards the applied mathematics and control communities. It is highly suggested that the authors revise the exposition of ideas and concepts, putting more emphasis on the power system so that it can be well-understood by the power system experts.}

{\nib \blue{TBD.}}

\noindent
{\bf R2.3. Comments by reviewer}:\\
{\em How will the proposed approach be used for the cases in which dependencies among uncertainty sources exist? An explanation and illustration would be helpful.}

{\nib \blue{In our paper, we do not make any assumption on statistical independence of the uncertainties. \ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}. Please note that the initial joint PDFs (28) and (30) are taken to be in that form for the sake of concrete example simulations. Their only purpose, in the simulations reported in Sec V, is to generate joint samples, and to evaluate the initial joint PDF values at the joint samples thus generated. If this initial scattered point cloud data is supplied to us only as a numerical table, without specifying any functional form whatsoever, nothing changes in the proposed algorithm. Similarly, the parameters in Table I could instead be correlated random vectors, and all we need are the \ul{joint} parametric weighted point cloud data. In fact, this is how we expect the proposed computational framework to be applied in practice.}}

@@

\noindent
{\bf R2.4. Comments by reviewer}:\\
{\em Incorporation of power system dynamics into the proposed uncertainty propagation framework is a bit vague. The proposed study seems not to have utilized any tool capable of simulating power system dynamics. Could the authors explain the part pertaining to the dynamic simulation in more detail?}

{\nib \blue{In this paper, we are interested in propagating the joint PDF values of the states of a power system dynamic model subject to joint stochastic uncertainties. Our focus is on explicitly computing these transient joint state PDFs at the time-varying state vector samples, which amounts to co-evolving the states and the joint PDF values over time. \ul{This is different from Monte Carlo or the traditional density estimation techniques since they approximate the joint PDFs via post-processing, not as an online computation}. To the best of our knowledge, there does not exist a power system simulation tool which can do this online computation while making neither statistical nor dynamical approximation.

Please note that in Sec. V, we have used MATPOWER and the Python-library ANDES (refs. [59] and [60]) for obtaining some parameters, as explained in the first paragraph of V-A. }}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R2.4. Comments by reviewer}:\\
{\em Incorporation of power system dynamics into the proposed uncertainty propagation framework is a bit vague. The proposed study seems not to have utilized any tool capable of simulating power system dynamics. Could the authors explain the part pertaining to the dynamic simulation in more detail?}

{\nib \blue{In this paper, we are interested in propagating the joint PDF values of the states of a power system dynamic model subject to joint stochastic uncertainties. Our focus is on explicitly computing these transient joint state PDFs at the time-varying state vector samples, which amounts to co-evolving the states and the joint PDF values over time. \ul{This is different from Monte Carlo or the traditional density estimation techniques since they approximate the joint PDFs via post-processing, not as an online computation}. To the best of our knowledge, there does not exist a power system simulation tool which can do this online computation while making neither statistical nor dynamical approximation.

Please note that in Sec. V, we have used MATPOWER and the Python-library ANDES (with inline citation of refs. [59] and [60] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

\noindent
{\bf R2.4. Comments by reviewer}:\\
{\em Incorporation of power system dynamics into the proposed uncertainty propagation framework is a bit vague. The proposed study seems not to have utilized any tool capable of simulating power system dynamics. Could the authors explain the part pertaining to the dynamic simulation in more detail?}

{\nib \blue{In this paper, we are interested in propagating the joint PDF values of the states of a power system dynamic model subject to joint stochastic uncertainties. Our focus is on explicitly computing these transient joint state PDFs at the time-varying state vector samples, which amounts to co-evolving the states and the joint PDF values over time. \ul{This is different from Monte Carlo or the traditional density estimation techniques since they approximate the joint PDFs via post-processing, not as an online computation}. To the best of our knowledge, there does not exist a power system simulation tool which can do this online computation while making neither statistical nor dynamical approximation.

Please note that in Sec. V, we have used MATPOWER and the Python-library ANDES (with inline citation of refs. [61] and [62] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}

@@

\noindent
{\bf R3.1. Comments by reviewer}:\\
{\em My main concern with this work is that all its derivations are deeply model-dependent and, unfortunately, the model that the author consider is not adequate to represent modern power systems. Moreover, second order model of synchronous machines are adequate only in a time scale of few seconds, so, even if the system were effectively including only conventional power plants, the model would be acceptable only for very short-term transients, i.e., transients for which the propagation of the noise is immaterial. In other words, the proposed model is useless.}

{\nib \blue{The reviewer's point is well taken. TBD.}}

\noindent
{\bf R3.2. Comments by reviewer}:\\
{\em The literature is incomplete, many recent works on SDAEs for power system applications are not mentioned.}

{\nib \blue{We appreciate the comment. As per the reviewer's suggestion, in the revised manuscript, we have included a sentence (marked blue) in Sec. I.1, second paragraph (last sentence) mentioning the usage of SDAEs in power system dynamic simulation. That sentence in the revised manuscript contains two newly added references: [22] and [23], both on simulating SDAE sample paths.}}

@@

\noindent
{\bf R3.3. Comments by reviewer}:\\
{\em There are several typos in the test (especially in the references).}

{\nib \blue{TBD. We corrected the two typos pointed out in comment \textbf{R4.8} in this rebuttal document.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
initiated TPS rebuttal version 1

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite-dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general.

In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal,geng2020reach} for set-valued analysis. These, however, require approximating the underlying nonlinear differential algebraic equations (DAEs) appearing in power system dynamic simulation, and thus lead to conservative analysis. For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets. Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as function of the uncertainties.

@@

\subsection{Statistical Mechanics Perspective}
A new difficulty for our SDE \eqref{ItoSDEvectorlevel} is that we have \emph{anisotropic} degenerate diffusion, i.e., the strengths of the noise acting in the last $n$ components of \eqref{ItoSDEvectorlevel} are nonuniform since $\bm{M}^{-1}\bm{\Sigma}$ is not identity. This complicates the matter because the construction of the functional $\Phi$ in \eqref{ProxRecursionInfiniteDim} is usually motivated via free energy considerations utilizing the structure of the \emph{stationary} PDF $\rho_{\infty}(\bm{\theta},\bm{\omega})$ for \eqref{KineticFPK}. The $\rho_{\infty}$ is, in turn, guaranteed to be a unique \emph{Boltzmann distribution} of the form\footnote{here $Z$ is a normalizing constant known as the ``partition function".}
\begin{subequations}
\begin{align}
\rho_{\infty}\left(\bm{\theta},\bm{\omega}\right) &= \dfrac{1}{Z}\exp\left(-\beta H\right), \\
H(\bm{\theta},\bm{\omega}) &:= V(\bm{\theta}) + \frac{1}{2}\langle\bm{\omega},\bm{M\omega}\rangle,	
\end{align}	
\label{BoltzmannPDF}
\end{subequations}
\emph{if and only if} the so-called \emph{Einstein relation} \cite{hernandez1989equilibrium,chen2015fast} holds:
\begin{align}
\bm{\Sigma}\bm{\Sigma}^{\top} = \beta^{-1}\left(\bm{\Gamma} + \bm{\Gamma}^{\top}\right) \quad\text{for some}\;\beta > 0.
\label{EinsteinRelation}	
\end{align}
In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} is equivalent to the proportionality constraint: $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.

@@

\begin{algorithm}[t]
\caption{Proposed proximal algorithm for $\tilde{\bm{\varrho}}_{k-1}\mapsto\tilde{\bm{\varrho}}_{k}$}
\label{algo:KineticProx}
\begin{algorithmic}[1]
\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
\For{$i=1$ to $N$}
\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-\left(\bm{\eta}_{k-1}^{i}\right)^{\top}\bm{\eta}_{k-1}^{i}-1\right)$
\For{$j=1$ to $N$}
\State $\bm{C}_{k}(i,j) \gets \parallel\bm{\eta}_{k-1}^{j} - \bm{\eta}_{k}^{i}+ h\nabla U(\bm{\xi}_{k}^{i})\parallel_{2}^{2}$
\State \qquad\quad $+12\bigg\|\dfrac{\bm{\xi}_{k-1}^{j}-\bm{\xi}_{k}^{i}}{h} - \dfrac{\bm{\eta}_{k-1}^{j} + \bm{\eta}_{k}^{i}}{2}\bigg\|_{2}^{2}$
\EndFor
\EndFor
\State $\bm{\Gamma}_{k} \gets \exp\left(-\bm{C}_{k}/2\varepsilon\right)$ \Comment{elementwise exponential}
\State $\bm{z}_{0} \gets {\rm{rand}}_{N\times 1}$ \Comment{random vector of size $N\times 1$}
\State $\bm{z} \gets \left[\bm{z}_{0}, \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\bm{y} \gets \left[\tilde{\bm{\varrho}}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\ell = 1$
\While{$\ell\leq \ell_{\max}$}
\State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
\State $\bm{y}(:,\ell+1) \gets \tilde{\bm{\varrho}}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
\If{$\|\bm{y}(:,\ell+1)-\bm{y}(:,\ell)\|_{2}<\delta$ \& $\|\bm{z}(:,\ell+1)-\bm{z}(:,\ell)\|_{2}<\delta$} \Comment{error within tolerance}
\State break
\Else
\State $\ell \gets \ell + 1$
\EndIf
\EndWhile\\
\Return $\tilde{\bm{\varrho}}_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
\EndProcedure	
\end{algorithmic}
\end{algorithm}

@@

>>>>

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite-dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general. {\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}. These studies, however, simulate the sample paths and does not directly propagate the joint state PDFs.}}

In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal,geng2020reach} for set-valued analysis. These, however, require approximating the underlying nonlinear (deterministic) DAEs appearing in power system dynamic simulation, and thus lead to conservative analysis. For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets. Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as function of the uncertainties.

@@

\subsection{Statistical Mechanics Perspective}
A new difficulty for our SDE \eqref{ItoSDEvectorlevel} is that we have \emph{anisotropic} degenerate diffusion, i.e., the strengths of the noise acting in the last $n$ components of \eqref{ItoSDEvectorlevel} are nonuniform since $\bm{M}^{-1}\bm{\Sigma}$ is not identity. This complicates the matter because the construction of the functional $\Phi$ in \eqref{ProxRecursionInfiniteDim} is usually motivated via free energy considerations utilizing the structure of the \emph{stationary} PDF $\rho_{\infty}(\bm{\theta},\bm{\omega})$ for \eqref{KineticFPK}. The $\rho_{\infty}$ is, in turn, guaranteed to be a unique \emph{Boltzmann distribution} of the form\footnote{here $Z$ is a normalizing constant known as the ``partition function".}
\begin{subequations}
\begin{align}
\rho_{\infty}\left(\bm{\theta},\bm{\omega}\right) &= \dfrac{1}{Z}\exp\left(-\beta H\right), \quad\text{for some}\;\beta > 0,\\
H(\bm{\theta},\bm{\omega}) &:= V(\bm{\theta}) + \frac{1}{2}\langle\bm{\omega},\bm{M\omega}\rangle,	
\end{align}	
\label{BoltzmannPDF}
\end{subequations}
\emph{if and only if} the so-called \emph{Einstein relation} \cite{hernandez1989equilibrium,chen2015fast} holds:
\begin{align}
\bm{\Sigma}\bm{\Sigma}^{\top} = \beta^{-1}\left(\bm{\Gamma} + \bm{\Gamma}^{\top}\right).
\label{EinsteinRelation}	
\end{align}
In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} is equivalent to the proportionality constraint: $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.

@@

\begin{algorithm}[t]
\caption{Proposed proximal algorithm for $\tilde{\bm{\varrho}}_{k-1}\mapsto\tilde{\bm{\varrho}}_{k}$}
\label{algo:KineticProx}
\begin{algorithmic}[1]
\Procedure{Prox}{$\tilde{\bm{\varrho}}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
\For{$i=1$ to $N$}
\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-\left(\bm{\eta}_{k-1}^{i}\right)^{\top}\bm{\eta}_{k-1}^{i}-1\right)$
\For{$j=1$ to $N$}
\State $\bm{C}_{k}(i,j) \gets \parallel\bm{\eta}_{k-1}^{j} - \bm{\eta}_{k}^{i}+ h\nabla U(\bm{\xi}_{k}^{i})\parallel_{2}^{2}$
\State \qquad\quad $+12\bigg\|\dfrac{\bm{\xi}_{k-1}^{j}-\bm{\xi}_{k}^{i}}{h} - \dfrac{\bm{\eta}_{k-1}^{j} + \bm{\eta}_{k}^{i}}{2}\bigg\|_{2}^{2}$
\EndFor
\EndFor
\State $\bm{\Gamma}_{k} \gets \exp\left(-\bm{C}_{k}/2\varepsilon\right)$ \Comment{elementwise exponential}
\State $\bm{z}_{0} \gets {\rm{rand}}_{N\times 1}$ \Comment{random vector of size $N\times 1$}
\State $\bm{z} \gets \left[\bm{z}_{0}, \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\bm{y} \gets \left[\tilde{\bm{\varrho}}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
\State $\ell = 1$
\While{$\ell\leq \ell_{\max}$}
\State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
\State $\bm{y}(:,\ell+1) \gets \tilde{\bm{\varrho}}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
\If{$\|\bm{y}(:,\ell+1)-\bm{y}(:,\ell)\|_{2}<\delta$ \& $\|\bm{z}(:,\ell+1)-\bm{z}(:,\ell)\|_{2}<\delta$} \Comment{error within tolerance}
\State break
\Else
\State $\ell \gets \ell + 1$
\EndIf
\EndWhile\\
\Return $\tilde{\bm{\varrho}}_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
\EndProcedure	
\end{algorithmic}
\end{algorithm}

@@

% === stochastic DAEs ======

\bibitem{wang2011numerical}
K.~Wang, and M.L.~Crow, ``Numerical simulation of stochastic differential algebraic equations for power system transient stability with random loads", \emph{2011 IEEE Power \& Energy Society General Meeting (PESGM)}, pp. 1--8, 2011. doi: 10.1109/PES.2011.6039188

\bibitem{milano2013systematic}
F.~Milano, and R.~Z{\'a}rate-Mi{\~n}ano, ``A systematic method to model power systems as stochastic differential algebraic equations", \emph{IEEE Transactions on Power Systems}, Vol. 28, No. 4, pp. 4537--4544, 2013.
